Dynamics of Laterally Propagating Flames in X-ray Bursts

### 2019-05-13 xrb/paper.tex
some more updates

\cite{cavecchi:2013} \MarginPar{and others}

@@

sponge\MarginPar{write}

@@

@@

To understand the time and length scales involved in flame
propagation, we do a 1D simulation of a laminar flame using our
microphysics.  Figure~\ref{fig:flame} shows the flame properties for a
He flame using the standard conductivities and our simple reaction
network.  This flame had a density of $2\times 10^6~\gcc$ and
temperature of $5\times 10^7$~K.  We note that this flame speed is
about $2\times 10^4~\cms$, the flame width is about 20~cm, and it
takes about 0.1~s to setting into a sustained flame.  Note this speed
is quite small compared to the speed estimated in
\citet{spitkovsky2002}.  We expected a multidimensional flame to
accelerate due to hydrodynamics interactions (wrinkling, turbulence
interactions, directed flows feeding fuel into the flame, ...).

We measure the flame width as:
\begin{equation}
\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
\end{equation}

To make the simulation feasible, we artifically speed up the flame.
Dimensional analysis shows that the flame speed scales as:\MarginPar{ref}
\begin{equation}
v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}}
\end{equation}
and the flame width scales as:
\begin{equation}
\lambda_f = \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
\end{equation}
To boost the flame, while keeping the thickness the same, we can
multiply both the burning rate and the conductivity by the same
factor.  We choose 10 for each, to give a flame 10$\times$ faster.

Experience with modeling resolved flames suggests that we need a
spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$.  These
conditions represent the bottom of the He layer.  As the density
decreases with altitude, the flame thickness increases and the flame
speed decreases, so we will easily resolve the flame structure throughout
the rest of the atmosphere.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{model}
\caption{\label{fig:sketch} A schematic of our model setup.}
\end{figure}

@@

We blend the hot and cold models laterally to produce the perturbation
needed to initiate a localized flame, with the hot model at the
origin of the axisymmetric geometry.  The blending is done as:
\begin{align}
p(x,y) &= f(x) p_\mathrm{hot}(y) + [1-f(x)] p_\mathrm{cool}(y) \\
\rho(x,y) &= f(x) \rho_\mathrm{hot}(y) + [1-f(x)] \rho_\mathrm{cool}(y) \\
X_k(x,y) &= f(x) {X_k}_\mathrm{hot}(y) + [1-f(x)] {X_k}_\mathrm{cool}(y)
\end{align}
with
\begin{equation}
f(x) = \begin{cases}
     1 & x < x_\mathrm{pert} \\
   1 - \frac{x - x_\mathrm{pert}}{\delta_\mathrm{blend}} & x_\mathrm{pert} \le x \le x_\mathrm{pert} + \delta_\mathrm{blend} \\
     0 & x > x_\mathrm{pert} + \delta_\mathrm{blend}
\end{cases}
\end{equation}
Since the equation of hydrostatic equilibrium is linear and our
blending is a linear combination of two models in hydrostatic
equilibrum, the blended model is also in equilibrium initially.  We choose
$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
the blended model is constructed, we compute $T(y)$ and $(\rho e)(y)$
from the equation of state.  \MarginPar{show}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{initial_model_paper}
\caption{\label{fig:initial_models} Our ``cool'' (solid) and ``hot''
  (dashed) initial models, showing both the density and temperature.}
\end{figure}

@@

\acknowledgements \castro\ is open-source and freely available at
\url{http://github.com/AMReX-Astro/Castro}.  The problem setups used
here are available in the git repo as {\tt flame} and {\tt
  flame\_wave}.  The work at Stony Brook was supported by DOE/Office
of Nuclear Physics grant DE-FG02-87ER40317.  This research used
resources of the National Energy Research Scientific Computing Center,
a DOE Office of Science User Facility supported by the Office of
Science of the U.~S.\ Department of Energy under Contract
No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
INCITE program.

@@

\software{ADS (http://adsabs.harvard.edu/),
          AMReX (\url{https://github.com/AMReX-Codes/amrex}),
          Castro \citep{castro},
          GCC (\url{https://gcc.gnu.org/}),
          linux (\url{https://www.kernel.org/}),
          matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/}),
          valgrind \citep{valgrind},
          yt \citep{yt}}

>>>>

\cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal} \MarginPar{need a summary}

@@

sponge\AssignTo{Mike will write}

@@

The speed and thickness of a laminar helium flame is determined by the
energy generation rate and conductivity and scale roughly as \MarginPar{need a ref}
\begin{equation}
v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}}
\end{equation}
and the flame width scales as:
\begin{equation}
\lambda_f = \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
\end{equation}
At the densities we consider in this simulation, the pure He flame
speed is quite slow and would require a prohibitively long integration
time to see significant evolution of the burning.  We therefore
consider boosted flames in this first paper, to accelerate the burning
and allow us to understand the qualitative effects of laterally
propagating flames.  To boost the flame, while keeping the thickness
the same, we can multiply both the burning rate and the conductivity
by the same factor.  For our standard calculations, we choose 10 for
each, to give a flame 10$\times$ faster.  We call this the ``10/10''
flame.  We will also do a simulation with the reactions and
conductivity both boosted by 5, the ``5/5'' flame.

To understand the time and length scales involved in flame
propagation, we do a 1D simulation of a laminar flame using our
microphysics.  Figure~\ref{fig:flame} shows the flame properties for a \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
He flame using the standard conductivities and our simple reaction
network.  This flame had a density of $2\times 10^6~\gcc$ and \MarginPar{double check}
temperature of $5\times 10^7$~K.  We note that this flame speed is
about $2\times 10^4~\cms$, the flame width is about 20~cm, and it
takes about 0.1~s to settle into a sustained flame.  Note this speed
is quite small compared to the speed estimated in
\citet{spitkovsky2002}.  We expected a multidimensional flame to
accelerate due to hydrodynamics interactions (wrinkling, turbulence
interactions, directed flows feeding fuel into the flame, ...).

We measure the flame width as:
\begin{equation}
\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
\end{equation}
Experience with modeling resolved flames suggests that we need a
spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$~\citep{SNld}.  These
conditions represent the bottom of the He layer.  As the density
decreases with altitude, the flame thickness increases and the flame
speed decreases, so we will easily resolve the flame structure throughout
the rest of the atmosphere.

Flame properties are show in Table XX \MarginPar{we'll redo the table with 10/10 and 5/5/ and unboosted flame}

@@

@@

\begin{figure}[t]
\centering
\plotone{model}
\caption{\label{fig:sketch} A schematic of our model setup.}
\end{figure}

@@

We blend the hot and cold models laterally to produce the perturbation
needed to initiate a localized flame, with the hot model at the
origin of the axisymmetric geometry.  The blending is done as:
\begin{align}
p(x,y) &= f(x) p_\mathrm{hot}(y) + [1-f(x)] p_\mathrm{cool}(y) \\
\rho(x,y) &= f(x) \rho_\mathrm{hot}(y) + [1-f(x)] \rho_\mathrm{cool}(y) \\
X_k(x,y) &= f(x) {X_k}_\mathrm{hot}(y) + [1-f(x)] {X_k}_\mathrm{cool}(y)
\end{align}
with
\begin{equation}
f(x) = \begin{cases}
     1 & x < x_\mathrm{pert} \\
   1 - \frac{x - x_\mathrm{pert}}{\delta_\mathrm{blend}} & x_\mathrm{pert} \le x \le x_\mathrm{pert} + \delta_\mathrm{blend} \\
     0 & x > x_\mathrm{pert} + \delta_\mathrm{blend}
\end{cases}
\end{equation}
Since the equation of hydrostatic equilibrium is linear and our
blending is a linear combination of two models in hydrostatic
equilibrum, the blended model is also in vertical equilibrium initially.  We choose
$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
the blended model is constructed, we compute $T(y)$ and $(\rho e)(y)$
from the equation of state.  \MarginPar{show}

\begin{figure}[t]
\centering
\plotone{initial_model_paper}
\caption{\label{fig:initial_models} Our ``cool'' (solid) and ``hot''
  (dashed) initial models, showing both the density and temperature.}
\end{figure}

@@

\acknowledgements \castro\ is open-source and freely available at
\url{http://github.com/AMReX-Astro/Castro}.  The problem setups used
here are available in the git repo as {\tt flame} and {\tt
  flame\_wave}.  The work at Stony Brook was supported by DOE/Office
of Nuclear Physics grant DE-FG02-87ER40317.  This research used
resources of the National Energy Research Scientific Computing Center,
a DOE Office of Science User Facility supported by the Office of
Science of the U.~S.\ Department of Energy under Contract
No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
INCITE program.  This research has made use of NASA's Astrophysics
Data System Bibliographic Services.

@@

\software{AMReX (\url{https://github.com/AMReX-Codes/amrex}),
          Castro \citep{castro},
          GCC (\url{https://gcc.gnu.org/}),
          linux (\url{https://www.kernel.org/}),
          matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/}),
          valgrind \citep{valgrind},
          VODE \citep{vode},
          yt \citep{yt}}
